Molecular dynamics simulation of synchronization in driven particles

### 2021-01-17 manuscript/manuscript_ajp_2020.tex
updated Fig.3, exercises

%Given the prevalence of synchronization,
%it is the subject of many computational models.
Here we perform numerical studies 
on the synchronized dynamics
of confined particles driven over
a washboard shaped potential energy landscape.
%Numerical studies %of harmonic systems
%can yield insights
%to the mechanics of naturally oscillating systems.
We chose this model for its
relevance to condensed matter systems
and ease of simulation.
%---------------------------------------------------------------
%kinks and nanofriction
%-----------------------------------------------------------------
%transition from synchronized to chaotic
%describe types of physical systems - particle types
%%In this work,
%we perform 
%numerical simulations of confined, driven particles
%to model a variety of physical phenomena.
%-----------------------------------------------------------------------
%paper outline
%In the following paper
In Section~\ref{sec:colloids} we describe colloid particles,
an relevent
experimental analog of our simulations.
We describe
our molecular dynamics model for a single particle in Section~\ref{sec:MD}.
%In this section, we restrict our discussion to a single particle,
%and include additional information
%for multiple particles in Section~\ref{sec:sync}.
%including various  and confining environments.
We summarize
our results 
%in Section~\ref{sec:results} 
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one}.
In Section~\ref{sec:sync}
we 
extend the model to include multiple particles,
and show how confined 
interacting particles respond to the same applied driving force.
%including stationary propagation of high density kinks
%in Sec.~\ref{sec:kink}.
%We include code to simulate
%and visualize the landscape and dynamics in this section
%and supplementary material.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
%We present these results using standard tools of non-linear oscillators
%such phase diagrams of velocity versus position.
%PHASE DIAGRAMS SHOWING COUPLING, OR QUASI-PERIODIC COUPLING
%INDICATE COOPERATIVE MOTION IN THE Q1D SUBSTRATE
%In Section~\ref{sec:quasiperiod},
%we show how an aperiodic landscape modifies the particle dynamics.
%Finally we explore the transition to chaotic dynamics in 
%Section~\ref{sec:chaos}
%and conclude 
In Section~\ref{sec:conclusion}
we describe how our work with colloids applies more broadly
to many physical systems
such as dusty plasmas, superconducting vortices and Josephson junctions.
We conclude with
exercises for interested students 
in Section~\ref{sec:problems}.

@@

%Here we model the microscopic dynamics of phase locking in a colloidal model syestem using molecular dynamics (MD).

@@

In Fig.~\ref{fig:0} 
the driven particle adjusts
its motion to traverse
the periodic landscape with the same rhythm as 
the applied time-dependent force $F^d(t)$.
%
In Fig.~\ref{fig:0}(a)
we plot $F^d(t)$ %defined in Eq.~\ref{eq:drive}.
as a function of time with
%The time independent
constants $F^{dc}=0.1$, 
$F^{ac}=0.05$ and $f=0.01$ cycles per time unit $\tau$.
The temporal period of the driving force is
$T = 1/f = 100 \tau$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%RAMP
%To avoid transient oscillations at early times, the dc driving force is initially zero and slowly increased %$F^{dc}$. at a rate of $\Delta F^{dc} = 0.01$ every $\Delta t = 4$ time units  to the constant value $F^{dc}=0.1$. In Fig.~\ref{fig:0}(a) we plot $F^{dc}$ in blue to show the increase in  $F^{dc}$ for the first $40$ time units In orange we show  the combined ac and dc drive $F^d(t)$ when the magnitude of the ac applied force is initially zero.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%since $\sin{(0)}=0$.
We mark the maxima of $F^{ac}(t)$
with the
vertical dashed lines %in panels (a) and (b)
-- i.e. $\sin{(\omega t)} = 1$.

%with the initial driving force at is minimum value 
%$F^d(0) = F^{dc}$.

In Fig.~\ref{fig:0}(b) 
we show the $y-$position of the particle
as a function of time.
We 
normalize $y$ by $\lambda$ %, the spatial period of the potential
to show how the particle moves between substrate minima.  
The initial particle position is $y = 0 $. %1 a_0$,
%where the $F^l$ is negative.
%JUST BECAUSE
%causing the particle to move to
%a landscape minima at the start of the simulation.
As the driving force increases,
%the effects of the landscape become less obvious,
the motion of the particle
synchronizes with the driving period $T$.
We measure 
average particle velocity $\langle{v}_y\rangle$
as the average displacement %$\Delta y = y(t_0+T) - y(t_0)$
over the period of the driving force.
The particle moves
in the positive y-direction
through $\Delta y = 2 \lambda$ over the time period $T$,
with 
the average velocity 
$\langle {v}_y \rangle= 2 \lambda f$. % where $T = 1/f$.
The horizontal dashed lines
coincide with the condition the substrate force is
minimum $F^l = -A_p$,  
i.e. $\sin{(\pi y / \lambda)} = 1$.
%or $\pi y / \lambda = m \pi/2$, where $m = 1, 5, 7$, {\it etc.}
%Once the particle motion
%coincides with the applied drive,
%indicating synchronization,
The %shows %indicates the
driving force is maximum when the landscape 
force on the particle is minimum,
as indicated with the coincidence of 
the particle position $y$
with the intersection of
the horizontal and vertical lines.
The slope of $y/\lambda$
is proportional to the net force on the particle.
This can be seen in 
Fig.~\ref{fig:0} %we demonstrate this relationship,
where local extrema appear in $y/\lambda$
when $F_{D}(t) \approx 0 $.

@@

The competition between the driving force and landscape potential
can produce a variety of hopping patterns in the particle motion. 
%In Fig.~\ref{fig:0}
%reaching a minimum position
%of $y/\lambda = 1$
%over the total period $T = 100$.
%In Fig. ~\ref{fig:0}
%the average displacement during one time period is a single wavelength
%of the substrate $y(t_0+T) - y(t_0) = \lambda$.
%single dc case with figure
%dc_current 0.01
%ac_current 0.05
%ac_frequency 0.01
%fp_y 0.1
%
The relative values of $F^{ac}$, $F^{dc}$ and $A_p$
control the rate and distance a  particle moves 
forward and backward in the landscape.
To explore the possible hopping patterns,
we sweep through a range of $F^{dc}$ for fixed $F^{ac}$ and $A_p$
to illustrate
a variety of oscillation modes as a function of $F^{dc}$.
A mode is a periodic pattern of hops
with a constant average particle velocity, $\langle {v}_{y} \rangle$
over a range of driving forces $F^{dc}$.
%as shown
In Fig.~\ref{fig:1} 
we keep $F^{dc}$ constant
for $10^5$ simulation timesteps
and 
measure the average velocity $\langle v_y \rangle $ 
as a function of $F^{dc}$,
with $F^{ac} = 0.05$, $f=0.01$, and the landscape %force $\vec{F}^{l}$
shown in Fig.~\ref{fig:0}.
%We plot the results in Fig.~\ref{fig:1} for 
The modes appear as 
discrete steps in $\langle v_y \rangle$ %occur because
across a range of $F^{dc}$,

%the average velocity of the particle
%is constant 
%despite the overall magnitude
%of $F^{dc}$.
Each step represents a different pattern of hops
between substrate minima
performed by the particle
due to the landscape confinement.  
In Fig.~\ref{fig:1} at %this occurs at 
low $F^{dc}$ the average velocity $\langle v_y \rangle$ is zero.
Since 
$A_p$ is large compared to the extrema of $F^{d}(t)$,
the particle oscillates back and forth
in a single minima with no net velocity.
At higher $F^{dc}$ the particle velocity 
$\langle v_{y} \rangle$ is increasing in steps of uniform height,
$\langle v_{y} \rangle = n \lambda f$,
where $n$ is an integer.
The step width is non-linear 
and can have a variety of
interesting patterns - including
a Devil's staircase pattern \cite{Bak1986}.
%$\lambda = L/N_p = 36.5/20 = 1.825$ is the spatial period, or wavelength
%of the landscape,
%and $f = 0.01$ cycles per time unit.

This synchronization
of the particle velocity can be measured
in terms of the landscape period.
Phase locked steps occur for average displacements integer
multiples of the substrate period $n\lambda$.

We explore a range of parameters in Ex.~\ref{ex:parameters}
where 
we perform this sweep for several values
of $F^{ac}$ including the limiting case that the applied force
is a constant value.
When $F^{ac} = 0$, no steps appear
in the $\langle v_y \rangle $ vs $F^{dc}$ curve. 
When $F^{dc} - F^{ac} < -A_p$ the particle
moves in the negative y-direction 
during part of the cycle.

@@

\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{sweep_FDC_vs_vx.pdf}
\caption{Average particle velocity  $\langle v_{y} \rangle$ as a function of $F^{dc}$,
  the constant parameter of $F^d(t)$ defined in Eq.~\ref{eq:drive}.
  The remaining parameters in $F^d(t)$ are as in Fig.~\ref{fig:0}, $F^{ac}=0.05$ and $f = 0.01$.}
\label{fig:1}
\end{figure}
\end{center}

@@

%We examine
The period of the substrate
can be used to control the
intrinsic velocity of the dc driven particle,
and the applied ac drive can cause mode-locking
which appears as non-linear 
steps in the force-velocity relationship.

Fundamental models
of synchronization 
can include many coupled oscillators
or focus on 
a single particle driven across a periodic potential landscape.
Single particle studies 
are useful to understand synchronization
in the absence of collective effects.
%A single particle particle oscillator
%within a potential well is 
%much like a skateboarder in a half-pipe or
%a child on a swing
%moving back and forth. % through equilibrium.  % \cite{}.
The relationship between the driving force and
landscape can change the pattern of motion.  
With a constant or 
dc drive the landscape modulates 
the particle velocity, below some threshold  
the dc force is not strong enough to push the particle
across a potential maximum so the average velocity is zero,
a phenomena refered to as pinning \cite{Reichhardt2017}.  
%and dissipating energy from the system at all driving forces.
A sinusoidal or ac drive creates mode-locking,
where the average particle velocity
is fixed for a range of dc drive forces \cite{Reichhardt2015}.

@@

\begin{enumerate}
\item Stokes' law describes the viscous drag on a sphere
  moving at velocity $\vec{v}$ as 
  $\vec{F}^{lin} = 3 \pi \eta D \vec{v}$ \cite{}
  where $\eta$ is the dynamic fluid viscosity and 
  $D$ is the particle diameter.
  (In simulation we
  subsume the constants $3 \pi D$
  such that $\eta \rightarrow 3 \pi D \eta$.)
  Often drag forces are
  modeled as a polynomial series
  $\vec{F}^{drag} = \vec{F}^{lin} + \vec{F}^{quad} + ... $.
  Truncating the series to the first term
  is justified by calculating the Reynolds number  
  $R = D v \rho / \eta$
  where $\rho$ is the fluid density and $v$ the particle's speed.
  When $R$ is small, the quadratic and higher order drag terms
  may be ignored in favor of the linear drag term.

  Demonstrate the Reynolds number is low
  for this system.

  Reasonable experimental values for
  a colloid are $D = 1 \mu m$ and $v \sim 1 \mu m /s$.
  Viscosity of water decreases with temperature
  from $1.3 \times 10^{-3} Pa-s$ at $10^{\circ}C$ to 
  from $0.31 \times 10^{-3} Pa-s$ at $90^{\circ}C$,
  making $\eta \sim 10^{-3} Pa-s$,
  where 1 pascal $ = 1 Pa = 1 N/m^2$.
  Likewise the density $\rho$ varies with temperature
  and $\rho \sim 10^3 kg/cm^3$ is a reasonable approximation.

  [citation for these values better than wikipedia] \cite{} %CONFIRM!
  %density $\rho = 0.9982 g/cm^3$
  %https://ascelibrary.org/doi/pdf/10.1061/9780784408230.ap02

This model should be familiar to readers
of a standard classical mechanics text, such as Ref.~\cite{Taylor2005}
when studying 
models Millikan's oil drops.

[could certainly do more with this question to make it more computational.  For instance, plot $R$ as a function of any of the parameters $v$, $D$ and $\rho$, include a temperature dependent $\eta(T)$ in the Brownian system.  Suggest students model something big and include inertia and quadratic drag, but then we're fairly off topic...]

@@

\item{Equation of Motion and Integration Methods}

  Newton's second law states that
  the acceleration of a particle $i$
  is proportional to 
  the sum of forces on a particle 
  \begin{equation}
  m_i \vec{a}_i = \sum \vec{F}_i
  \label{eq:n2l}
  \end{equation}
  where the constant of proportionality is the
  inertial mass $m_i$
  The addition of a dissipative force in a dynamical model
  on a colloid is typically modeled
  with a drag force proportional to the particle's velocity
  in the opposite direction of motion 
  $\vec{F}^{drag} = - \eta \vec{v}_i$
  where $\eta$ is the drag coefficient.
  Demonstrate
  the ratio of $m/\eta$ has units of time, 
  %This ratio of the acceleration coefficient $m$
  %to the drag coefficient $\eta$
  %is
  known as the momentum relaxation time.
  The momentum relaxation time is 
  known to be small for
  particles with low Reynolds numbers,
  indivating 
  a regime in which the acceleration term can be ignored
  entirely.
  Demonstrate that a particle confined to a landscape exerting force
  $F^l(\vec{r}_i)$ subject to a time dependent drive $F^d(t)$
  can be modeled with the equation of motion described in 
  Eq.~\ref{eq:motion}. 
  %\label{ex:n2l}

%\item
  Generally when solving differential equations,
  the Euler method is effective for solving linear equations,
  such as the one-dimensional 
  $f(y) = df/dy$ 
  but not higher order methods.

  The Verlet method simplifies to

@@

\item Drawing the contour

  The code for generating
a two dimensional colored plot
of the potential landscape
is calculated by evaluating
the analytic function in Eq.~\ref{eq:landscape}
for a grid of values $(x_n,y_n)$.
%is included in the supplementary material,

@@

\item
  \label{ex:parameters}

\item Frenkel-Kontorova and competing length scales for low temperature systems where thermal fluctuations are negligible

  \item The Fokker-Planck Equation

  \item Aubrey-Kontorova + commensurability

  \item nanotribology

    kink motion
    Vanossi et al J.Phys.: Condens. Matter 19 (2007)

@@

\section{Conclusion}
\label{sec:conclusion}	% You can label sections for reference
Our simulations reproduce results presented in 
Juniper {\it et al.} \cite{Juniper2015, Juniper2018}
which demonstrated
mode locking in
experiments and simulations of 
driven colloids on a
optical periodic landscape.
Due to the relative ease of
manipulation and imagining, 
colloids are used as experimental models
for systems relatively hard to access and visualize. 
%SAME SENTENCE TWICE
Collections of colloids can
be used to study the properties of solids and liquids
%comprised of smaller particles
%such as atoms \cite{} or electrons \cite{}
or more exotic systems 
such as cold atoms or electron gases \cite{Grier2003}.

@@

    \bibitem{Stewart1968}  W. C. Stewart CURRENT‐VOLTAGE CHARACTERISTICS OF JOSEPHSON JUNCTIONS Appl. Phys. Lett. 12, 277 (1968).

@@

@@

>>>>

%Given the prevalence of synchronization,
%it is the subject of many computational models.
Here we perform numerical studies 
on the synchronized dynamics
of confined particles driven over
a washboard shaped potential energy landscape.
%Numerical studies %of harmonic systems
%can yield insights
%to the mechanics of naturally oscillating systems.
We chose this model for its
relevance to condensed matter systems
and ease of simulation.
%---------------------------------------------------------------
%kinks and nanofriction
%-----------------------------------------------------------------
%transition from synchronized to chaotic
%describe types of physical systems - particle types
%%In this work,
%we perform 
%numerical simulations of confined, driven particles
%to model a variety of physical phenomena.
%-----------------------------------------------------------------------
%paper outline
%In the following paper
In Section~\ref{sec:colloids} we describe colloid particles,
an relevant
experimental analog of our simulations.
We describe
our molecular dynamics model for a single particle in Section~\ref{sec:MD}.
%In this section, we restrict our discussion to a single particle,
%and include additional information
%for multiple particles in Section~\ref{sec:sync}.
%including various  and confining environments.
We summarize
our results 
%in Section~\ref{sec:results} 
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one}.
In Section~\ref{sec:sync}
we 
extend the model to include multiple particles,
and show how confined 
interacting particles respond to the same applied driving force.
%including stationary propagation of high density kinks
%in Sec.~\ref{sec:kink}.
%We include code to simulate
%and visualize the landscape and dynamics in this section
%and supplementary material.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
%We present these results using standard tools of non-linear oscillators
%such phase diagrams of velocity versus position.
%PHASE DIAGRAMS SHOWING COUPLING, OR QUASI-PERIODIC COUPLING
%INDICATE COOPERATIVE MOTION IN THE Q1D SUBSTRATE
%In Section~\ref{sec:quasiperiod},
%we show how an aperiodic landscape modifies the particle dynamics.
%Finally we explore the transition to chaotic dynamics in 
%Section~\ref{sec:chaos}
%and conclude 
In Section~\ref{sec:conclusion}
we describe how our work with colloids applies more broadly
to many physical systems
such as dusty plasmas, superconducting vortices and Josephson junctions.
We conclude with
exercises for interested students 
in Section~\ref{sec:problems}.

@@

%Here we model the microscopic dynamics of phase locking in a colloidal model system using molecular dynamics (MD).

@@

%In Fig.~\ref{fig:0} 
The driven particle adjusts
its motion to traverse
the periodic landscape with the same rhythm as 
the applied time-dependent force $F^d(t)$.
%
In Fig.~\ref{fig:0}(a)
we plot $F^d(t)$ %defined in Eq.~\ref{eq:drive}.
as a function of time with
%The time independent
constants $F^{dc}=0.1$, 
$F^{ac}=0.05$ and $f=0.01$ cycles per time unit $\tau$.
The temporal period of the driving force is
$T = 1/f = 100 \tau$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%RAMP
%To avoid transient oscillations at early times, the dc driving force is initially zero and slowly increased %$F^{dc}$. at a rate of $\Delta F^{dc} = 0.01$ every $\Delta t = 4$ time units  to the constant value $F^{dc}=0.1$. In Fig.~\ref{fig:0}(a) we plot $F^{dc}$ in blue to show the increase in  $F^{dc}$ for the first $40$ time units In orange we show  the combined ac and dc drive $F^d(t)$ when the magnitude of the ac applied force is initially zero.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%since $\sin{(0)}=0$.
We mark the maxima of $F^{ac}(t)$
with the
vertical dashed lines %in panels (a) and (b)
-- i.e. $\sin{(\omega t)} = 1$.
%
%
%with the initial driving force at is minimum value 
%$F^d(0) = F^{dc}$. 
%
In Fig.~\ref{fig:0}(b) 
we show the $y-$position of the particle
as a function of time.
We 
normalize $y$ by $\lambda$ %, the spatial period of the potential
to show  particle motion between substrate minima.  
The initial particle position is $y = 0 $. %1 a_0$,
%where the $F^l$ is negative.
%JUST BECAUSE
%causing the particle to move to
%a landscape minima at the start of the simulation.
As the driving force increases,
%the effects of the landscape become less obvious,
the motion of the particle
synchronizes with the driving period $T$.
We measure 
average particle velocity $\langle{v}_y\rangle$
as the average displacement %$\Delta y = y(t_0+T) - y(t_0)$
over the period of the driving force.
The particle moves
in the positive y-direction
through $\Delta y = 2 \lambda$ over the time period $T$,
with 
the average velocity 
$\langle {v}_y \rangle= 2 \lambda f$. % where $T = 1/f$.
The horizontal dashed lines
coincide with %the condition the is
minimum substrate force  $F^l = -A_p$,  
i.e. $\sin{(\pi y / \lambda)} = 1$.
%or $\pi y / \lambda = m \pi/2$, where $m = 1, 5, 7$, {\it etc.}
%Once the particle motion
%coincides with the applied drive,
%indicating synchronization,
The particle synchronizes its motion such that the 
driving force is maximum when the landscape 
force is minimum,
as indicated with %the coincidence of 
with the intersection of
horizontal and vertical lines
and the particle position $y$.
The slope of $y/\lambda$
is proportional to the net force on the particle.
This can be seen in 
Fig.~\ref{fig:0} %we demonstrate this relationship,
where local extrema appear in $y/\lambda$
when $F_{D}(t) \approx 0 $.

@@

The competition between the driving force and landscape potential
can produce a variety of hopping patterns in the particle motion. 
%In Fig.~\ref{fig:0}
%reaching a minimum position
%of $y/\lambda = 1$
%over the total period $T = 100$.
%In Fig. ~\ref{fig:0}
%the average displacement during one time period is a single wavelength
%of the substrate $y(t_0+T) - y(t_0) = \lambda$.
%single dc case with figure
%dc_current 0.01
%ac_current 0.05
%ac_frequency 0.01
%fp_y 0.1
%
The relative values of $F^{ac}$, $F^{dc}$ and $A_p$
control the rate and distance a  particle moves 
forward and backward in the landscape.
To explore the possible hopping patterns,
we sweep through a range of $F^{dc}$ for fixed $F^{ac}$ and $A_p$
to illustrate
distinct oscillation modes as a function of $F^{dc}$.
A mode is a periodic pattern of hops
with a constant average particle velocity, $\langle {v}_{y} \rangle$
over a range of driving forces $F^{dc}$.
%as shown
In Fig.~\ref{fig:1} 
we keep $F^{dc}$ constant
for $10^5$ simulation timesteps
and 
measure the average velocity $\langle v_y \rangle $ 
as a function of $F^{dc}$,
with $F^{ac} = 0.05$, $f=0.01$, and the landscape %force $\vec{F}^{l}$
shown in Fig.~\ref{fig:0}.
%We plot the results in Fig.~\ref{fig:1} for 
The modes appear as 
discrete steps in $\langle v_y \rangle$ %occur because
across a range of $F^{dc}$,

%the average velocity of the particle
%is constant 
%despite the overall magnitude
%of $F^{dc}$.
Each step represents a different pattern of hops
between substrate minima
performed by the particle
due to the landscape confinement.  
In Fig.~\ref{fig:1} at %this occurs at 
low $F^{dc}$ the average velocity $\langle v_y \rangle$ is zero.
Since 
$A_p$ is large compared to the extrema of $F^{d}(t)$,
the particle oscillates back and forth
in a single minima with no net velocity.
At higher $F^{dc}$ the particle velocity 
$\langle v_{y} \rangle$ increases in steps of uniform height,
$\langle v_{y} \rangle = n \lambda f$,
where $n$ is an integer.
The step width is non-linear 
and can have a variety of
interesting patterns \cite{Bak1986}.
%$\lambda = L/N_p = 36.5/20 = 1.825$ is the spatial period, or wavelength
%of the landscape,
%and $f = 0.01$ cycles per time unit.

%This synchronization
%of the particle velocity can be measured
%in terms of the landscape period.
%Phase locked steps occur for average displacements integer
%multiples of the substrate period $n\lambda$.

We explore a range of parameters in Ex.~\ref{ex:parameters}
where 
we perform this sweep for several values
of $F^{ac}$ including the limiting case that the applied force
is a constant value.
When $F^{ac} = 0$, no steps appear
in the $\langle v_y \rangle $ vs $F^{dc}$ curve. 
When $F^{dc} - F^{ac} < -A_p$ the particle
moves in the negative y-direction 
during part of the cycle.
%We examine
The period of the substrate
can be used to control the
intrinsic velocity of the dc driven particle,
and the applied ac drive can cause mode-locking
which appears as non-linear 
steps in the force-velocity relationship.

@@

\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{sweep_FDC_vs_vx.pdf}
\caption{Average particle velocity  $\langle v_{y} \rangle$ as a function of $F^{dc}$,
  the constant parameter of $F^d(t)$ defined in Eq.~\ref{eq:drive}.
  The remaining parameters in $F^d(t)$ are as in Fig.~\ref{fig:0}, $F^{ac}=0.05$ and $f = 0.01$. [{\it I believe the slight slope in this curve is due to sampling - where does the particle start/end ?}]}
\label{fig:1}
\end{figure}
\end{center}

@@

[Insert Phase Plots section/figure?]

@@

\begin{enumerate}
\item {\bf Drag Models and Reynolds Numbers}
  Stokes' law describes the viscous drag on a sphere
  moving at velocity $\vec{v}$ as 
  $\vec{F}^{lin} = -3 \pi \eta D \vec{v}$ 
  where $\eta$ is the dynamic fluid viscosity and 
  $D$ is the particle diameter \cite{Taylor2005}.
  In simulation we
  subsume the constants $3 \pi D$
  such that $\eta \rightarrow 3 \pi D \eta$.
  Often drag forces are
  modeled as a polynomial series
  $\vec{F}^{drag} = -b \vec{v} - c v^2 \hat{v} + ... $.
  Truncating the series to the first term
  is justified by demonstrating the sphere
  has a low Reynolds number  
  $R = D v \rho / \eta$
  where $\rho$ is the fluid density and $v$ the particle's speed.
  When $R$ is small, the quadratic and higher order drag terms
  may be ignored in favor of the linear drag term.
  %
  Using reasonable values for the
  experimental analog of this system, 
  demonstrate the Reynolds number is low
  for this system.

  Reasonable experimental values for
  a colloid are $D = 1 \mu m$ and $v \sim 1 \mu m /s$.
  Viscosity of water decreases with temperature
  from $1.3 \times 10^{-3}$ Pa-s at $10^{\circ}$ C to 
  from $0.31 \times 10^{-3}$ Pa-s at $90^{\circ}$ C,
  making $\eta \sim 10^{-3}$ Pa-s,
  where 1 Pascal is $1$ N/m$^2$.
  Likewise the density $\rho$ varies with temperature
  and $\rho \sim 10^3$ kg/cm$^3$ is a reasonable approximation.
  \cite{asce}

  This model should be familiar to readers
  of a standard classical mechanics text, such as Ref.~\cite{Taylor2005}
  when studying 
  models Millikan's oil drops.
  [{\it could certainly do more with this question to make it more computational.  For instance, plot $R$ as a function of any of the parameters $v$, $D$ and $\rho$, include a temperature dependent $\eta(T)$ in the Brownian system.  Suggest students model something big and include inertia and quadratic drag, but then we're fairly off topic...}]

@@

\item{\bf Equation of Motion}

  Newton's second law states that
  the acceleration of a particle $i$
  is proportional to 
  the sum of forces on a particle 
  \begin{equation}
  m_i \vec{a}_i = \sum \vec{F}_i
  \label{eq:n2l}
  \end{equation}
  where the constant of proportionality is the
  inertial mass $m_i$
  The addition of a dissipative force in a dynamical model
  on a colloid is typically modeled
  with a drag force proportional to the particle's velocity
  in the opposite direction of motion 
  $\vec{F}^{drag} = - \eta \vec{v}_i$
  where $\eta$ is the drag coefficient.
  The ratio of $m/\eta$ has units of time, 
  %This ratio of the acceleration coefficient $m$
  %to the drag coefficient $\eta$
  %is
  known as the momentum relaxation time.
  The momentum relaxation time is 
  known to be small for
  particles with low Reynolds numbers,
  indicating 
  a regime in which the acceleration term can be ignored
  entirely.
  (a) Using Newton's Second Law for
  a small momentum relaxation time, 
  demonstrate that a particle confined to a landscape exerting force
  $F^l(\vec{r}_i)$ subject to a time dependent drive $F^d(t)$
  can be modeled with the equation of motion described in 
  Eq.~\ref{eq:motion}. 
  \label{ex:n2l}

  \item {\bf Integration Methods}
  Our equation of motion provides
  a direct calculation for particle velocity.
  To calculate the position of the particle we
  integrate the equation of motion using
  the standard definition of velocity
  $\vec{v}_i = d\vec{r}_i/dt$.
  %
  In our model we use the
  Euler method to solve this equation.
  Generally when solving ordinary differential equations,
  the Euler method is effective for solving linear equations
  of the form
  $dy/dt = f(t,y(t))$ with initial condition $y(t_0) = y_0$.
  The solution is calculated algorithmically
  by stepping in time through $n$ integer steps
  $t_n = t_0 + n \Delta t$.
  At each subsequent step the new
  value for $y$ is calculated as a map solution using
  discrete times 
  $y_{n+1} = y_n + f(t_n,y_n)$.
  The Euler method can be applied to non-linear
  equations effectively if the time step $\Delta t$
  is kept sufficiently small.  
  \cite{Newman}

  Generally molecular dynamics algorithms
  are solved with higher order methods
  such as the Verlet method or Runge-Kutta methods.

@@

\item {\bf Visualizing the motion}
  Drawing the contour

  The code for generating
a two dimensional colored plot
of the potential landscape
is calculated by evaluating
the analytic function in Eq.~\ref{eq:ysubstrate}
for a grid of values $(x_n,y_n)$.
%is included in the supplementary material,

Also the animation library in python...

@@

\item {\bf Exploring Parameters of the Model}
  \label{ex:parameters}

%\item Frenkel-Kontorova and competing length scales for low temperature systems where thermal fluctuations are negligible
%  \item The Fokker-Planck Equation
%  \item Aubrey-Kontorova + commensurability
%  \item nanotribology
 %   kink motion
 %   Vanossi et al J.Phys.: Condens. Matter 19 (2007)

@@

\section{Conclusion}
\label{sec:conclusion}	% You can label sections for reference
Fundamental models
of synchronization 
can include many coupled oscillators
or focus on 
a single particle driven across a periodic potential landscape.
Single particle studies 
are useful to understand synchronization
in the absence of collective effects.
%A single particle particle oscillator
%within a potential well is 
%much like a skateboarder in a half-pipe or
%a child on a swing
%moving back and forth. % through equilibrium.  % \cite{}.
The relationship between the driving force and
landscape can change the pattern of motion.  
With a constant or 
dc drive the landscape modulates 
the particle velocity, below some threshold  
the dc force is not strong enough to push the particle
across a potential maximum so the average velocity is zero,
a phenomena referred to as pinning \cite{Reichhardt2017}.  
%and dissipating energy from the system at all driving forces.
A sinusoidal or ac drive creates mode-locking,
where the average particle velocity
is fixed for a range of dc drive forces \cite{Reichhardt2015}.

Our simulations reproduce results presented in 
Juniper {\it et al.} \cite{Juniper2015, Juniper2018}
which demonstrated
mode locking in
experiments and simulations of 
driven colloids on a
optical periodic landscape.
Due to the relative ease of
manipulation and imagining, 
colloids are used as experimental models
for systems relatively hard to access and visualize. 
%SAME SENTENCE TWICE
Collections of colloids can
be used to study the properties of solids and liquids
%comprised of smaller particles
%such as atoms \cite{} or electrons \cite{}
or more exotic systems 
such as cold atoms or electron gases \cite{Grier2003}.

@@

    \bibitem{Stewart1968}  W. C. Stewart Current‐Voltage Characteristics of Josephson Junctions, Appl. Phys. Lett. 12, 277 (1968).

@@

      \bibitem{Newman} M. Newman, Computational Physics...

@@

\bibitem{asce}
  What is the method for citing non-fundamental constants?
    (1) https://ascelibrary.org/doi/pdf/10.1061/9780784408230.ap02
    (2) IAPWS 2008 - http://www.iapws.org/relguide/viscosity.html
  %[citation for these values better than wikipedia] \cite{} %CONFIRM!
  %density $\rho = 0.9982 g/cm^3$
